---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?
A1. The most frequent donation is between $10 and $25, and for the most part, the smaller the donation category, the more frequent the donation.

```{r}
wv_winred_contribs_cateogries <- wv_winred_contribs %>% 
#select(last_name, first_name, amount) %>% 
mutate(
 amount_category = case_when(
  amount < 10 ~ "Less than $10",
  amount >= 10 & amount <= 25 ~ "Between $10 and $25",
  amount >= 26 & amount <= 50 ~ "Between $26 and $50",
  amount >= 51 & amount <= 100 ~ "Between $51 and $100",
  amount >= 101 & amount <= 500 ~ "Between $101 and $500",
  amount >= 501 ~ "$501 and up"
 )
) %>% 
group_by(amount_category) %>% 
  summarise(
    count_contribs = n()
  ) %>%
  arrange(desc(count_contribs))
```

Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. 

```{r}
wv_winred_contribs_ten <- wv_winred_contribs %>% 
  filter(amount < 10) %>% 
  group_by(last_name, first_name) %>% 
  summarise(
    count = n()
  ) %>% 
  arrange(desc(count))
```

Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. My biggest question based on the results of Question 2 are, who is Reverend Dr. John Almond and why has he made 293 separate donations of 10 dollars or less? Why wouldn't Almond and the other frequent contributors of 10 dollars or less just make one large donation? Did he/others regularly donate based on factors like when he got paid? Did he have a schedule for donating? Did he just throw in some money whenever he felt like it? I want to sort Almond's contributions by date to get to the bottom of this.

Q4. Which month saw the most contributions?
A4. April saw the most contributions of the three months in the data set.

```{r}
wv_winred_contribs_month <- wv_winred_contribs %>%
  mutate(month = month(date)) %>%
  group_by(month) %>% 
  mutate(
    sum_amount = sum(amount)
  ) %>% 
  select(month, sum_amount) %>% 
  distinct() %>% 
  arrange(desc(sum_amount))
```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.
A5. There were six contributions of $2900, three of which went to CAROL FOR CONGRESS, the other three each went to different congressional campaigns. Carol Miller is from West Virginia, as is David Mckinley. The other two recipients are from Montana and Mississippi, respectively.

```{r}
wv_winred_contribs_max <- wv_winred_contribs %>%
 filter(amount == 2900)
```
