---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(lubridate)
library(tidyverse)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
nonvoters_data <- read_csv("data/nonvoters_data.csv")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Nonvoters were more likely to have lower incomes; to be young; to have lower levels of education; and to say they don’t belong to either political party, which are all traits that square with what we know about people less likely to engage with the political system.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
nonvoters_categories <- nonvoters_data %>% 
  mutate(
    Q30 = case_when(
      Q30 >= 3 ~ "neither political party",
      Q30 == 1 ~ "Republican",
      Q30 == 2 ~ "Democrat",
    )) %>% 
  mutate(
    ppage = case_when(
      ppage <= 24 ~ "18-24",
      ppage > 24 ~ "older than 24")
  )
nonvoters_categories %>% 
  summarize(income_cat, educ, Q30, voter_category, ppage)
```
### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-