---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(lubridate)
library(tidyverse)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
nonvoters_data <- read_csv("data/nonvoters_data.csv")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Nonvoters were more likely to have lower incomes; to be young; to have lower levels of education; and to say they don’t belong to either political party, which are all traits that square with what we know about people less likely to engage with the political system.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
nonvoters_categories <- nonvoters_data %>%
  mutate(
    Q30 = case_when(
      Q30 >= 3 ~ "neither political party",
      Q30 == 1 ~ "Republican",
      Q30 == 2 ~ "Democrat",
      )) %>%
  mutate(
    ppage = case_when(
      ppage <= 24 ~ "22-24",
      ppage > 24 ~"older than 24") )
nonvoters_categories %>%
  summarize(income_cat, educ, Q30, voter_category, ppage)
```
```{r}
nonvoters_total <- nonvoters_categories %>%
  group_by(voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n())
nonvoters_total
```
```{r}
total_nonvoters <- 1451
```

```{r}
nonvoters_party <- nonvoters_categories %>%
  group_by(Q30,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
    mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_party
```

```{r}
nonvoters_income <- nonvoters_categories %>%
  group_by(income_cat,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_income
```

```{r}
nonvoters_age <- nonvoters_categories %>%
  group_by(ppage,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_age
```

```{r}
nonvoters_edu <- nonvoters_categories %>%
  group_by(educ,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_edu
```
### Sentence 2

* **Sentence text**: Our survey found, for instance, that occasional voters were slightly more likely than frequent voters to have a long-term disability (8 percent vs. 5 percent), and nonvoters were even more likely to fall into this category (12 percent). Black and Hispanic voters are also more likely to experience hurdles, perhaps in part because there tend to be fewer polling places in their neighborhoods. About 24 percent of Black respondents said that they had to stand in line for more than an hour while voting, and Hispanic respondents were more likely to say they had trouble accessing the polling place or couldn’t get off work in time to vote.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
disability_voters
```
```{r}
# Put code to reverse engineer sentence here
never_disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
never_disability_voters
```

```{r}
sporadic_voters_total <- 2574
always_voters_total <- 1811
```


```{r}
# Put code to reverse engineer sentence here
always_disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/always_voters_total)*100) %>%
arrange(desc(percent))
always_disability_voters
```
```{r}
# Put code to reverse engineer sentence here
sporadic_disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sporadic_voters_total)*100) %>%
arrange(desc(percent))
sporadic_disability_voters
```

# Display results of code below this codeblock

```{r}
black_voters <- nonvoters_data %>%
  group_by(race) %>%
  filter(race == "Black") %>%
  summarise(count = n())
black_voters <- 932
```

```{r}
line_voters <- nonvoters_data %>%
  group_by(Q18_8,race) %>%
  filter(race == "Black") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/black_voters)*100) %>%
arrange(desc(percent))
line_voters
```

```
```


### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: This year, some of these occasional voters are also dealing with pandemic-specific challenges that could make voting less of a priority. For instance, people who vote more irregularly are a little more likely than those who almost always vote to have lost a job in the past year (13 percent vs. 10 percent) or to have worried about expenses (21 percent vs. 16 percent), although those who almost never vote are even likelier than those who sometimes vote to have reported these experiences.

* **Analysis summary**: In our analysis, we found 

```{r}
never_votes_job <- nonvoters_data %>%
  group_by(Q11_1,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
never_votes_job
```
```{r}
sporadic_votes_job <- nonvoters_data %>%
  group_by(Q11_1,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
sporadic_votes_job
```
```{r}
always_votes_job <- nonvoters_data %>%
  group_by(Q11_1,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
always_votes_job
```
```{r}
never_votes_worry <- nonvoters_data %>%
  group_by(Q11_5,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
never_votes_worry
```
```{r}
sporadic_votes_worry <- nonvoters_data %>%
  group_by(Q11_5,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
sporadic_votes_worry
```
```{r}
always_votes_worry <- nonvoters_data %>%
  group_by(Q11_5,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
always_votes_worry
```



### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-