------
title: "Data Analysis"
author: "Abby Wallace, Lisa Woelfl, Torrence Banks, Noah Ferguson"
date: "11/11/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are analyzing data from the Maryland Lobbying Registrations. [https://lobby-ethics.maryland.gov/]

## Load libraries

```{r echo=FALSE, message=FALSE}
# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
```

## Load and Cleaning Data

```{r}
# Loading the dataframes and binding them together
one <- read_csv("registrations(1).csv") %>% clean_names()
two <- read_csv("registrations(2).csv") %>% clean_names()
three <- read_csv("registrations(3).csv") %>% clean_names()
four<- read_csv("registrations(4).csv") %>% clean_names()
five <- read_csv("registrations(5).csv")%>% clean_names()
six <- read_csv("registrations(6).csv") %>% clean_names()
seven <- read_csv("registrations(7).csv") %>% clean_names()
eight <- read_csv("registrations(8).csv") %>% clean_names()
nine <- read_csv("registrations(9).csv") %>% clean_names()
total_registrations <- rbind(one, two, three, four, five, six, seven, eight, nine)
write_csv(total_registrations, "total_registrations.csv")
# We cleaned the data in OpenRefine and split the date column
clean_total_registrations <- read_csv("clean_total_registrations.csv")
clean_total_registrations[c('start', 'end')] <- str_split_fixed(clean_total_registrations$registration_period, '-', 2)
```

## Basic explorations

```{r}
glimpse(clean_total_registrations)
```

The original dataframe had 12855 rows and 5 columns. Our cleaned dataframe has 12855 rows and 9 columns (we added cleaned columns for organization name & employer and split the date column to have a start and end date for the registration).

The data looked pretty clean in OpenRefine. We found just a handful of names to clean up.

Limitations: The dataframe is pretty basic with a small number of columns. That means we will have to cross-reference the website, if we want to know how much a lobbyist is making or how much an employer is spending on lobbying. The standard dataset doesn't even include the registration matters. For our topical questions, we have to download new datasets for every registration matter we're interested in. Another limitation is that the earliest data is from 2019, which means we can't look at long-term trends.

To answer most of our questions, we will have to do additional research. For example, we can ask the number of employers and lobbyists in the energy sector and see how that changed over the last few years. To examine how much the top firms make, we have to go back to the website.

When looking into the PG county youth curfew, we won't get clear cut answers from the data alone. Trying to find a connection between the number of lobbyists for matters of juvenile law and the implementation of the curfew will require additional research on our part.

### Question 1

* **Question text**: What lobbyist/registrant represents the most employers in our data set?

* **Analysis summary**: The lobbyist that represents the most employers in our data set was Timothy A. Perry. He's been representing employers for 12 years as a member of Perry White Ross Jacobson LLC and served as Chief of Staff to Senate President Thomas V. Mike Miller, Jr., the longest serving State Senate President in United States history. He lives in Baltimore County, Maryland, and has represented 126 distinct employers in our joined data set. He's represented employers such as Stevenson University and AT&T over four of our data cycles. According to a profile written about him in September 2022, he specializes in government relations and serves as a board member of First Fruits Farm, a charitable nonprofit organization that specializes in food services. (https://marquistoplawyers.com/2022/09/16/timothy-perry/)

In the below block, I counted the number of total employers that each lobbyist represents, then arranged in descending order to see who has the most total employers in the cleaned data set.

```{r}
most_employers <- clean_total_registrations %>% 
  group_by(lobbyist_registrant) %>% 
  count() %>% 
  arrange(desc(n))
```

Here, I filtered for Timothy A. Perry's different employers across each of the data sets to see the different employers he represents. There, I realized that he's represented a lot of the employers over multiple data sets, effectively skewing the above data.

```{r}
clean_total_registrations %>% 
  filter(lobbyist_registrant == "Perry, Timothy  A"
  )
```

I used the codeblock below to see which of Perry's employers he's represented over the most datasets. For example, he's represented Anheuser-Busch Companies four different times, which makes the 333 number in the most_employers data set no longer the number I want to see. I want to see how many unique companies each lobbyist represented, and the current code accounts for all of the duplicates instead of each distinct employer.

```{r}
dupe_employers <- clean_total_registrations %>% 
   filter(lobbyist_registrant == "Perry, Timothy  A") %>% 
  count(clean_employer, source = TRUE) %>% 
  arrange(desc(n))
```

By grouping by lobbyist, organization firm and employer and counting for distinct employers before ungrouping and regrouping by lobbyist, I was able to figure out how many unique employers each lobbyist represents. Perry still leads the way but with a far smaller number than the original data set after all of the duplicate employers were removed. This number better answers the original question that we took from the data, which was who represents the most UNIQUE employers in the combined data set

```{r}
distinct_employers <- clean_total_registrations %>% 
    group_by(lobbyist_registrant, clean_organization_firm, clean_employer) %>% 
  summarise(
    count= n_distinct(lobbyist_registrant, clean_organization_firm, clean_employer)
  ) %>% 
  ungroup() %>%
  group_by(lobbyist_registrant) %>% 
  count() %>% 
arrange(desc(n))
```
