---
title: "pre_lab_01.Rmd"
author: "Derek Willis"
date: "2022-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## About this notebook

This notebook contains code and explanatory text that your should review and run as you read through chapters 1 - 5 of the course textbook, "Data Journalism with R and the Tidyverse". Answer questions and edit the document as directed.

Running this notebook will help you understand key data analysis methods and concepts that you will put into practice during this week's lab. When you come across a new command or verb in R, you should add it to your class reference notebook file along with what it's used for.

When you are finished running the code in this notebook, you will push changes to your course GitHub repo, and upload the link to ELMS as instructed.

You also should update your class reference notebook with code examples and what they mean. When you come across a new concept or R verb, put it in your reference notebook along with a sentence about what it does. For example:

```{r}
library(tidyverse)
```

The library function loads an R library such as the tidyverse, providing access to the verbs and functions that library contains. You have to load a library before you can use its functions.

## Instructor in class demonstration outline

* Load this notebook, and provide overview of R, RStudio (text editor, console, environment window, file explorer), Rmarkdown data notebooks, the reference notebook.
* Work through the code below, chapter by chapter.
  * Chapter 3: explain the console and the concept of variables.
  * Chapter 3: explain packages, the Tidyverse, how to install them.
  * Chapter 6: explain loading libraries, the concept of functions, loading data, the importance of having questions to guide data analysis, and aggregation and summarization.

## Chapter 3

### Task 1: The console
**Task** In the console (look below), type 2+2 and hit enter. What does it return? Type your answer below.
**Answer:**

4 

### Task 2: Variables
**Task** Put the following code -- which stores the numeral 2 as a variable called number -- in the console and hit enter.    

```{r}
number <- 2
```

Then type this code to multiply the variable called number by itself.  

```{r}
number * number
```

What does it return? Type your answer below:

4
**Answer:**

### Task 3: Install packages
**Task** In the console, run the following code to install the Tidyverse:

```{r}
install.packages('tidyverse')
```

**Task** Add that

Then run the following code to install the rmarkdown library

```{r}
install.packages('rmarkdown')
```

## Chapter 6

### Task 1: Load libraries
**Task** Run the following code in the gray-colored codeblock below -- not in the console -- to load the tidyverse library. To run the code, click the little green play button (left facing arrow) at the top right of the codeblock. In Rmarkdown data notebooks, we write code inside of codeblocks, and explanatory text in the white area outside of it.   

```{r}
# This is a comment inside of a codeblock. Comments doesn't run when we run the codeblock.
# Adding this command is a good way to start your class reference notebook.
library(tidyverse)
```

### Task 2: Load data
**Task** Load the Maryland WinRed contributions dataset by running the following codeblock.

```{r}
maryland_winred_contributions <- read_rds("pre_lab_01.rds")
```

Once it loads, look at the environment window at right in RStudio for maryland_winred_contributions. It's next to a small blue circle with a white arrow in it.  To the right of that, look at how many "obs" the dataset has ("obs" is short for observation, which is another way of saying rows or records). Look at how many variables it has (another way of saying columns or fields).  Write the number of observations and variables below.

54247 observations and 24 variables
### Task 3: Glimpse data
**Task** Run the following codeblock to use glimpse() to get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.

```{r}
glimpse(maryland_winred_contributions)
```


In the space below, write the data type for the "name" column and the "amount" column (it's in light gray, offset by < >).  Also offer your best guess as to what the "name" column and the "amount" column represent.
<chr> is the data type for first and last name, and <dbl> is the data type for amount. My best guess is that the name column represents the first and last names of Maryland contributors, and the amount column represents how much they contributed.

### Task 4: Examine data "head"
**Task** Run the following codeblock to use head() to examine the first six rows of data. Explore the data by clicking the black arrow on the right side of the output table, allowing you to examine additional columns.
```{r}
head(maryland_winred_contributions)
```
In the space below, write the name of the business from the first row of the "name" column.
Beth Hansher, the first contributor on the row, is a Homemaker. It says "Information Requested" for her business.

### Task 5: Explore data with interactive dataframe
**Task** look at the environment window at right in RStudio for maryland_winred_contributions It's next to a small blue circle with a white arrow in it. Click on the maryland_winred_contributions -- not the blue circle. It will open the dataframe as an interactive table.  Scroll down to see additional rows.  Scroll left to see additional columns. Click on a column name to sort by that column.

### Task 6: Group and count
**Task** Run the following codeblock:

```{r}
maryland_winred_contributions %>%
  group_by(date) %>%
  summarise(
    count_contribs = n()
  )
```

Describe what you think this code is doing below.

This code is sorting contributions daily by date starting on April 1, 2022.

### Task 7
**Task** Run the following codeblock:

```{r}
maryland_winred_contributions %>%
  group_by(date) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange(desc(count_contribs))
```
Compared to the previous codeblock, what has changed and how did that affect the result?
The dates are no longer the column being sorted. It is now the contribution count, from largest to smallest, that is being counted regardless of date.

### Task 8
**Task** Run the following codeblock.
```{r}
maryland_winred_contributions %>%
  group_by(memo_text, amount) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange(desc(count_contribs))
```

Compared to the previous codeblock, what has changed and how did that affect the result?
The code is now showing which donations were most prevalent/frequent, sorting from highest to lowest the amount of contributions. For instance, 1419 people gave 25 dollars to SAVE AMERICA JOINT FUNDRAISING COMMITTEE, the moist frequent of any donation.

### Task 9
**Task** Run the following codeblock:
```{r}
maryland_winred_contributions %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount)
  ) %>%
  arrange(desc(total_amount))
```
Compared to the previous codeblock, what has changed and how did that affect the result?
The code is now sorting by the total amount donated instead of the most frequent type of donation. For instance, the NICOLEE AMBROSE FOR CONGRESS earmark only had 150 contributions, but rackex up 67,980 dollars, more than the row underneath it which had more total contributions.

### Task 10
**Task** Run the following codeblock to group the maryland_winred_contributions dataframe by the recipient for each contribution. Then do the following summarizations for each group:count the number of contributions, add up the value of all contributions, and calculate the mean and median contribution amounts. Then sort so the recipient with the most contributions is listed first.
```{r}
maryland_winred_contributions %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount)
  ) %>%
  arrange(desc(count_contribs))
```
Compared to the previous codeblock, what has changed and how did that affect the result? Glancing at the names, are any of the top results from Maryland?
The code is now sorting by number of contributions while showing the total, mean and median amounts for that campaign group. None of the top results were from Maryland.

### Task 11
**Task** Run the following codeblock to group the maryland_winred_contributions dataframe by the city for each contribution. Then do the following summarizations for each group:count the number of contributions, add up the value of all contributions, calculate the mean and median contributions amounts and calculate the smallest and largest contributions Then sort so the city with the largest total contribution amount is listed first.

```{r}
maryland_winred_contributions %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))
```
Compared to the previous codeblock, what has changed and how did that affect the result? How is the data sorted?
The data is now sorted highest to lowest based on total amount donated by city. Bethesda is at the top of the data set because its $139,757.20 was the most donated by any city in Maryland. The way that the data is displayed also shows how much on average each person is donating per city. Chevy Chase has a much higher average amount donated per donation than Bethesda, but has much fewer total contributors than Bethesda.

