---
title: "lab_07"
author: "derek willis"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already

```{r echo=FALSE}
census_api_key("d02ca188f0878d4314ebc4b1379d7aaa7e592844", overwrite=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2020 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2020 5-year ACS and save them to a dataframe

```{r}
acs2020 <- load_variables(2020, "acs5", cache = TRUE)
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2020 5-year estimates? You will need to find the variable for median age first.
**A1**The jurisdiction with the lowest median age is Baltimore City, with a median age of 35.5. 

```{r}
md <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2020) %>% 
  arrange(estimate)
```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50.
**A2** The two counties with median ages of at least 50 are Talbot County and Worcester County.

```{r}
md_50 <- md %>% 
  filter(
    estimate >= "50.0")
```

**Q3** We're interested in knowing which Maryland counties have the highest percentage of 5-17 year-olds who speak Spanish at home and *also* speak English fluently. Start with the code you wrote for Q1 and alter it to show the total population of 5-17 year-olds by county. Then create a second dataframe to show the number of 5-17 year-olds by county living in homes where Spanish is spoken but who speak English "very well" according to the ACS. To do this, you will need to find the ACS variables that represent those subjects. HINT: the phrase "very well" shows up in multiple variables - you want the one that doesn't include other factors like citizenship. 

For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.

Which county has the highest percentage of 5-17 year-olds who speak Spanish at home and also speak English very well?

**A3** Prince George's County has the highest percentage of 5-17 year-olds who speak Spanish at home and also speak English very well.

```{r}
md_five <- get_acs(geography = "county",
              variables = c(totalkids = "B16007_002"),
              state = "MD",
              year = 2020) %>% 
  select(GEOID, NAME, estimate) %>% 
  rename(totalkids = estimate)
```


```{r}
md_spanish <- get_acs(geography = "county",
              variables = c(spanish_english = "B16004_005"),
              state = "MD",
              year = 2020) %>% 
  select(GEOID, NAME, estimate) %>% 
  rename(spanish_english = estimate)
```

```{r}
md_combined <- md_spanish %>% 
 inner_join(md_five)
```

```{r}
md_percent <- md_combined %>% 
  mutate(
    percent = (spanish_english/totalkids)) %>% 
      arrange(desc(percent))
```



**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, which county stands out to you in the data as different from the others? How would you describe the top 5 counties? How would you describe Garrett County's numbers?

**A4** The top two counties (Montgomery County and Prince George's County) have large populations, and there's a steep drop off from 2 to 3. Talbot County has a large number of Spanish speakers despite being a much smaller county than some of the larger percentage counties. Garrett County's percentage is super low, which is probably because Garrett County is overwhelmingly white.

**Q5** Pennsylvanians are currently requesting and submitting mail ballots for the November election. You'd like to know more about the people who live in the areas sending in mail ballots. Read in the `zip_totals.csv` file in the data folder containing the number of returned mail ballots per zip code. Then create a dataframe of Pennsylvania ZCTAs that include the following variables from the 2019 ACS: total population and median income. You will need to add `output = "wide"` to the get_acs function so that you get columns for those two variables' estimates and margins of error.

Join the ZCTA dataframe to the zip totals dataframe and calculate a per capita rate for mail ballots per 100 population in each ZCTA, and write code to answer the following questions?

* Which zip code *that has returned at least 100 mail ballots* has the highest per capita rate? 
* Where is that zip code and what does Census Reporter say about its demographics? 
* For zip codes with a per capita rate of at least 10 returned ballots per 100 people, how many have median incomes of at least $100,000? 
* How many have median incomes less than the state figure of $68,957? 
* What's the most interesting one from this last result?

**A5**
1. 	Zip Code 19436 has the highest per capita rate of any code that has returned at least 100 mail ballots.
2. 19436 is Gwynedd, Pennsylvania. The population is primarily white. The number of seniors is extremely large while the number of people in their late 20s to early 40s is extremely small. There are also an extremely large number of single adults and an extremely small number of single parents. The percentage of children under 18 living in the 19436 ZIP code is extremely small compared to other areas of the country.
3. Zero zip codes with a per capita rate of at least 10 returned ballots per 100 people have median incomes of at least $100,000.
4. 1 zip code (15142) with a per capita rate of at least 10 returned ballots per 100 people has a median incomes of at least $100,000.
5. Zip code 15142 is made up of over 90% white people.

```{r}
pa_zip_totals <- read_csv("data/pa_zip_totals.csv")
```

```{r}
pa_zcta <- get_acs(geography = "zcta",
              variables = c(total_pop = "B01001_001", median_income = "B06011_001"),
              state = "PA",
              year = 2019, output = "wide")
```

```{r}
pa_zips_and_zcta <- pa_zip_totals %>% 
  inner_join(pa_zcta, by=c("zip" = "GEOID")) %>% 
  mutate(
    per_capita_100 = ((count/total_popE)*100)
  )
```

```{r}
at_least_100 <- pa_zips_and_zcta %>% 
  filter(count >= 100) %>% 
  arrange(desc(per_capita_100))
```

```{r}
at_least_10 <- pa_zips_and_zcta %>% 
  filter(per_capita_100 >= 10 & median_incomeE >= 100000)
```

```{r}
at_least_avg <- pa_zips_and_zcta %>% 
  filter(per_capita_100 >= 10 & median_incomeE >= 68957)
```
